Q1: Are lake and terrestrial primary productivity coherent?

### 2019-06-05 AnalysisCode/ms1_analyses.Rmd
ms1 result updates by region

```{r compute coherences, echo=TRUE, cache=TRUE}

@@

#agriculture -- is 500m buffer best? Other options include 100m buffer (probably too small) and hu12 watershed
pct.ag<-lagosne_select(table="buffer500m.lulc", vars=c("lagoslakeid","buffer500m_nlcd2001_pct_82","buffer500m_nlcd2006_pct_82","buffer500m_nlcd2011_pct_82"))
# pct.ag<-lagosne_select(table="hu12.lulc", vars=c("lagoslakeid","hu12_nlcd2001_pct_82","hu12_nlcd2006_pct_82","hu12_nlcd2011_pct_82"))
pct.ag<-pct.ag[pct.ag$lagoslakeid %in% analysislakes$lakeinfo$lagoslakeid,]
pct.ag.avg<-data.frame(lagoslakeid=pct.ag$lagoslakeid, pct.ag=rowMeans(pct.ag[,2:4]))

@@

predictors<-analysislakes$lakeinfo
predictors$tslength<-predictors$end-predictors$start+1
predictors<-left_join(predictors, depth, by="lagoslakeid")
predictors<-left_join(predictors, pct.ag.avg, by="lagoslakeid")
predictors<-left_join(predictors, avg.chla, by="lagoslakeid")
predictors<-left_join(predictors, tsi.chl, by="lagoslakeid")
predictors<-left_join(predictors, huc_codes, by="hu4_zoneid")

@@

modvars.accndvi.phist<-modvars.accndvi[modvars.accndvi$accndvip.ts1<0.2,]
modvars.accndvi.philt<-modvars.accndvi[modvars.accndvi$accndvip.ts2<0.2,]
```

```{r gls modelling, echo=TRUE, cache=FALSE}
#short timescales
gls.coh.accndvi.st<-gls(accndvicoh.ts1 ~ maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi,
                       correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(gls.coh.accndvi.st)
suppressWarnings(dredge.coh.accndvi.st<-dredge(gls.coh.accndvi.st, beta="sd")) #intercept only is best model. Disappointing.
print(head(dredge.coh.accndvi.st))

gls.p.accndvi.st<-gls(accndvip.ts1 ~ tslength + maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi,
                       correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(gls.p.accndvi.st)
suppressWarnings(dredge.p.accndvi.st<-dredge(gls.p.accndvi.st, beta="sd")) #intercept only is best model. Disappointing.
print(head(dredge.p.accndvi.st))

gls.phi.accndvi.st<-gls(cos(accndviphi.ts1) ~ maxdepth + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi.phist) #remove ftype b/c only lakes
summary(gls.phi.accndvi.st)
suppressWarnings(dredge.phi.accndvi.st<-dredge(gls.phi.accndvi.st, beta="sd")) #intercept only is best model. Disappointing.
print(head(dredge.phi.accndvi.st))

#long timescales
gls.coh.accndvi.lt<-gls(accndvicoh.ts2 ~ maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi,
                       correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(gls.coh.accndvi.lt)
suppressWarnings(dredge.coh.accndvi.lt<-dredge(gls.coh.accndvi.lt, beta="sd")) #intercept only is best model. Disappointing.
print(head(dredge.coh.accndvi.lt))

gls.p.accndvi.lt<-gls(accndvip.ts2 ~ tslength + maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi,
                       correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(gls.p.accndvi.lt)
suppressWarnings(dredge.p.accndvi.lt<-dredge(gls.p.accndvi.lt, beta="sd")) #intercept only is best model. Disappointing.
print(head(dredge.p.accndvi.lt))

gls.phi.accndvi.lt<-gls(cos(accndviphi.ts2) ~ maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi.philt,
                       correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(gls.phi.accndvi.lt)
suppressWarnings(dredge.phi.accndvi.lt<-dredge(gls.phi.accndvi.lt, beta="sd")) #intercept only is best model. Disappointing.
print(head(dredge.phi.accndvi.lt))
```

GLS models sucked, but inspection of results suggested that mixed effects models with random effects of region on the intercept may likely be an improvement.

```{r mixed modelling, echo=TRUE, cache=FALSE}
lme.coh.accndvi.st<-lmer(accndvicoh.ts1 ~ maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + (1|huc2_code), data=modvars.accndvi, na.action="na.fail")
summary(lme.coh.accndvi.st)
suppressWarnings(dredge.coh.accndvi.st<-dredge(lme.coh.accndvi.st, beta="sd")) #intercept only is best model. Disappointing.
print(head(dredge.coh.accndvi.st))

lme.p.accndvi.st<-lmer(accndvip.ts1 ~ tslength + maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + (1|huc2_code), data=modvars.accndvi, na.action="na.fail")
summary(lme.p.accndvi.st)
suppressWarnings(dredge.p.accndvi.st<-dredge(lme.p.accndvi.st, beta="sd")) #intercept only is best model. Disappointing.
print(head(dredge.p.accndvi.st))

lme.phi.accndvi.st<-lmer(accndviphi.ts1 ~ maxdepth + lake_area_ha + pct.ag + chla + tsi.cat + (1|huc2_code), data=modvars.accndvi.phist, na.action="na.fail")
summary(lme.phi.accndvi.st)
suppressWarnings(dredge.phi.accndvi.st<-dredge(lme.phi.accndvi.st, beta="sd")) #model with TSI is unconvincingly better than intercept-only (dAIC = 0.92)
print(head(dredge.phi.accndvi.st))

lme.coh.accndvi.lt<-lmer(accndvicoh.ts2 ~ maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + (1|huc2_code), data=modvars.accndvi, na.action="na.fail")
summary(lme.coh.accndvi.lt)
suppressWarnings(dredge.coh.accndvi.lt<-dredge(lme.coh.accndvi.lt, beta="sd")) #lake type is marginal improvement over intercept only (dAIC = 0.1)
print(head(dredge.coh.accndvi.lt))

lme.p.accndvi.lt<-lmer(accndvip.ts2 ~ tslength + maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + (1|huc2_code), data=modvars.accndvi, na.action="na.fail")
summary(lme.p.accndvi.lt)
suppressWarnings(dredge.p.accndvi.lt<-dredge(lme.p.accndvi.lt, beta="sd")) #intercept only is best model. Disappointing.
print(head(dredge.p.accndvi.lt))

lme.phi.accndvi.lt<-lmer(accndviphi.ts2 ~ maxdepth + lake_area_ha + pct.ag + chla + tsi.cat + (1|huc2_code), data=modvars.accndvi.philt, na.action="na.fail")
summary(lme.phi.accndvi.lt)
suppressWarnings(dredge.phi.accndvi.lt<-dredge(lme.phi.accndvi.lt, beta="sd")) #model with TSI is unconvincingly better than intercept-only (dAIC = 0.92)
print(head(dredge.phi.accndvi.lt))

@@

Adding random effects of huc2 region on intercept helped somewhat. Some possibilities are to add more detailed lake type classes, and to model 2-3 regions separately.

>>>>

```{r compute coherences, echo=TRUE, cache=FALSE}

@@

#agriculture -- is 500m buffer best? Other options include 100m buffer (probably too small) and hu12 watershed
# pct.ag<-lagosne_select(table="buffer500m.lulc", vars=c("lagoslakeid","buffer500m_nlcd2001_pct_82","buffer500m_nlcd2006_pct_82","buffer500m_nlcd2011_pct_82"))
pct.ag<-lagosne_select(table="hu12.lulc", vars=c("hu12_zoneid","hu12_nlcd2001_pct_82","hu12_nlcd2006_pct_82","hu12_nlcd2011_pct_82"))
pct.ag<-pct.ag[pct.ag$hu12_zoneid %in% analysislakes$lakeinfo$hu12_zoneid,]
pct.ag.avg<-data.frame(hu12_zoneid=pct.ag$hu12_zoneid, pct.ag=rowMeans(pct.ag[,2:4]))

@@

#state info
states<-lagosne_select(table="state", vars=c("state_zoneid","state_name"))

predictors<-analysislakes$lakeinfo
predictors$tslength<-predictors$end-predictors$start+1
predictors<-left_join(predictors, depth, by="lagoslakeid")
predictors<-left_join(predictors, pct.ag.avg, by="hu12_zoneid")
predictors<-left_join(predictors, avg.chla, by="lagoslakeid")
predictors<-left_join(predictors, tsi.chl, by="lagoslakeid")
predictors<-left_join(predictors, states, by="state_zoneid")
#predictors<-left_join(predictors, huc_codes, by="hu4_zoneid")

for(nn in 1:ncol(predictors)){

  if(is.factor(predictors[,nn])){
    predictors[,nn]<-factor(predictors[,nn])
  }

str(predictors)
hist(predictors$lake_area_ha)
hist(log10(predictors$lake_area_ha))
hist(predictors$lake_perim_meters)
hist(log10(predictors$lake_perim_meters))
table(predictors$nhd_fcode)
table(predictors$hu12_zoneid)
hist(predictors$elevation_m)
hist(predictors$maxdepth)
hist(log10(predictors$maxdepth))
hist(predictors$pct.ag)
hist(log10(predictors$pct.ag))
hist(predictors$chla)
hist(log10(predictors$chla))
hist(predictors$tsi)
table(predictors$tsi.cat)

predictors$log10_lake_area_ha<-log10(predictors$lake_area_ha)
predictors$log10_lake_perim_meters<-log10(predictors$lake_perim_meters)
predictors$log10_maxdepth<-log10(predictors$maxdepth)
predictors$log10_pct.ag<-log10(predictors$pct.ag+1)
predictors$log10_chla<-log10(predictors$chla)

@@

modvars.accndvi.phist<-modvars.accndvi[modvars.accndvi$accndvip.ts1<0.2,]
modvars.accndvi.philt<-modvars.accndvi[modvars.accndvi$accndvip.ts2<0.2,]

@@

```{r gls modelling, echo=TRUE, cache=FALSE}
#short timescales
gls.coh.accndvi.st<-gls(accndvicoh.ts1 ~ log10_maxdepth + log10_lake_area_ha + log10_pct.ag + log10_chla + tsi.cat, 
                        data=modvars.accndvi,
                        correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(gls.coh.accndvi.st)
suppressWarnings(dredge.coh.accndvi.st<-dredge(gls.coh.accndvi.st, beta="sd")) #intercept only is best model. Disappointing.
print(head(dredge.coh.accndvi.st))

gls.phi.accndvi.st<-gls(cos(accndviphi.ts1) ~ log10_maxdepth + log10_lake_area_ha + log10_pct.ag + log10_chla + tsi.cat, 
                        data=modvars.accndvi.phist,
                        correlation=corExp(form = ~ nhd_lat + nhd_long)) #remove ftype b/c only lakes
summary(gls.phi.accndvi.st)
suppressWarnings(dredge.phi.accndvi.st<-dredge(gls.phi.accndvi.st, beta="sd")) #intercept only is best model. Disappointing.
print(head(dredge.phi.accndvi.st))

#long timescales
gls.coh.accndvi.lt<-gls(accndvicoh.ts2 ~ log10_maxdepth + log10_lake_area_ha + log10_pct.ag + log10_chla + tsi.cat, 
                        data=modvars.accndvi,
                        correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(gls.coh.accndvi.lt)
suppressWarnings(dredge.coh.accndvi.lt<-dredge(gls.coh.accndvi.lt, beta="sd")) #intercept only is best model. Disappointing.
print(head(dredge.coh.accndvi.lt))

gls.phi.accndvi.lt<-gls(cos(accndviphi.ts2) ~ log10_maxdepth + log10_lake_area_ha + log10_pct.ag + log10_chla + tsi.cat, 
                        data=modvars.accndvi.philt,
                        correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(gls.phi.accndvi.lt)
suppressWarnings(dredge.phi.accndvi.lt<-dredge(gls.phi.accndvi.lt, beta="sd")) #intercept only is best model. Disappointing.
print(head(dredge.phi.accndvi.lt))
```
That turned up astonishingly little, so now try breaking down by region. 
The regions are: 1) Missouri; 2) Minnesota, Wisconsin, and Michigan; 3) Pennsylvania, New York, Rhode Island, Vermont, New Hampshire, Maine.

```{r gls modeling r1, echo=TRUE, cache=FALSE}
#Region 1: Missouri

modvars.accndvi.r1<-modvars.accndvi[modvars.accndvi$state_name=="Missouri",]
# modvars.accndvi.philt.r1<-modvars.accndvi.philt[modvars.accndvi.philt$state_name=="Missouri",]
# modvars.accndvi.phist.r1<-modvars.accndvi.phist[modvars.accndvi.phist$state_name=="Missouri",]

#short timescales
gls.coh.accndvi.st<-gls(accndvicoh.ts1 ~ log10_maxdepth + log10_lake_area_ha + log10_pct.ag + log10_chla + tsi.cat, 
                        data=modvars.accndvi.r1,
                        correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(gls.coh.accndvi.st)
suppressWarnings(dredge.coh.accndvi.st<-dredge(gls.coh.accndvi.st, beta="sd")) #intercept only is best model. Disappointing.
print(head(dredge.coh.accndvi.st))

#long timescales
gls.coh.accndvi.lt<-gls(accndvicoh.ts2 ~ log10_maxdepth + log10_lake_area_ha + log10_pct.ag + log10_chla + tsi.cat, 
                        data=modvars.accndvi.r1,
                        correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(gls.coh.accndvi.lt)
suppressWarnings(dredge.coh.accndvi.lt<-dredge(gls.coh.accndvi.lt, beta="sd")) #intercept only is best model. Disappointing.
print(head(dredge.coh.accndvi.lt))

```{r gls modeling r2, echo=TRUE, cache=FALSE}
#Region 2: Minnesota, Wisconsin, Michigan

modvars.accndvi.r2<-modvars.accndvi[modvars.accndvi$state_name %in% c("Minnesota","Wisconsin","Michigan"),]

#short timescales
gls.coh.accndvi.st<-gls(accndvicoh.ts1 ~ log10_maxdepth + log10_lake_area_ha + log10_pct.ag + log10_chla + tsi.cat, 
                        data=modvars.accndvi.r2,
                        correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(gls.coh.accndvi.st)
suppressWarnings(dredge.coh.accndvi.st<-dredge(gls.coh.accndvi.st, beta="sd")) #intercept only is best model. Disappointing.
print(head(dredge.coh.accndvi.st))

#long timescales
gls.coh.accndvi.lt<-gls(accndvicoh.ts2 ~ log10_maxdepth + log10_lake_area_ha + log10_pct.ag + log10_chla + tsi.cat, 
                        data=modvars.accndvi.r2,
                        correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(gls.coh.accndvi.lt)
suppressWarnings(dredge.coh.accndvi.lt<-dredge(gls.coh.accndvi.lt, beta="sd")) #intercept only is best model. Disappointing.
print(head(dredge.coh.accndvi.lt))

```
```{r gls modeling r3, echo=TRUE, cache=FALSE}
#Region 2: Minnesota, Wisconsin, Michigan

modvars.accndvi.r3<-modvars.accndvi[modvars.accndvi$state_name %in% c("Pennsylvania","New York","Rhode Island","Vermont","New Hampshire","Maine"),]

#short timescales
gls.coh.accndvi.st<-gls(accndvicoh.ts1 ~ log10_maxdepth + log10_lake_area_ha + log10_pct.ag + log10_chla + tsi.cat, 
                        data=modvars.accndvi.r3,
                        correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(gls.coh.accndvi.st)
suppressWarnings(dredge.coh.accndvi.st<-dredge(gls.coh.accndvi.st, beta="sd")) #intercept only is best model. Disappointing.
print(head(dredge.coh.accndvi.st))

#long timescales
gls.coh.accndvi.lt<-gls(accndvicoh.ts2 ~ log10_maxdepth + log10_lake_area_ha + log10_pct.ag + log10_chla + tsi.cat, 
                        data=modvars.accndvi.r3,
                        correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(gls.coh.accndvi.lt)
suppressWarnings(dredge.coh.accndvi.lt<-dredge(gls.coh.accndvi.lt, beta="sd")) #intercept only is best model. Disappointing.
print(head(dredge.coh.accndvi.lt))

### 2019-06-05 AnalysisCode/ms1_prep.Rmd
rough think piece display items
ms1 result updates by region

@@

lagoslakes.raw<-makeLAGOSannualts(lakes_to_get, infovars=c("lake_area_ha", "lake_perim_meters", "nhd_ftype", "hu4_zoneid"), tsvars=c("chla"), 
                                  minmos=3, minobs=3, timespan=c(1989,2018))
lagoslakes.cln<-cleanAnnualts(lagoslakes.raw)

@@

MNMPCAlakes.raw<-makeMNMPCAannualts(dir="/Users/jonathanwalter/Box Sync/NSF EAGER Synchrony/Data/LAGOS Extended/MN_MPCA",
                                    infovars=c("lake_area_ha", "lake_perim_meters", "nhd_ftype", "hu4_zoneid"), minmos=3, minobs=3, timespan=c(1989,2018))
MNMPCAlakes.raw<-fixNamestoLAGOS(MNMPCAlakes.raw,"Minnesota MPCA")
MNMPCAlakes.cln<-cleanAnnualts(MNMPCAlakes.raw)

WIDNRlakes.raw<-makeWIDNRannualts(dir="/Users/jonathanwalter/Box Sync/NSF EAGER Synchrony/Data/LAGOS Extended/WI_DNR",
                                  infovars=c("lake_area_ha", "lake_perim_meters", "nhd_ftype", "hu4_zoneid"), minmos=3, minobs=3, timespan=c(1989,2018))
WIDNRlakes.raw<-fixNamestoLAGOS(WIDNRlakes.raw, "Wisconsin DNR")
WIDNRlakes.cln<-cleanAnnualts(WIDNRlakes.raw)

>>>>

##updated cleanAnnualts on 5/5/19

@@

lagoslakes.raw<-makeLAGOSannualts(lakes_to_get, infovars=c("lake_area_ha", "lake_perim_meters", "nhd_ftype", "nhd_fcode", "hu4_zoneid", 
                                                           "hu12_zoneid", "state_zoneid", "elevation_m"), tsvars=c("chla"), 
                                  minmos=3, minobs=3, timespan=c(1989,2018))
lagoslakes.cln<-cleanAnnualts(lagoslakes.raw)

@@

MNMPCAlakes.raw<-makeMNMPCAannualts(dir="/Users/jonathanwalter/Box Sync/NSF EAGER Synchrony/Data/LAGOS Extended/MN_MPCA",
                                   infovars=c("lake_area_ha", "lake_perim_meters", "nhd_ftype", "nhd_fcode", "hu4_zoneid","hu12_zoneid", 
                                              "state_zoneid", 
                                              "elevation_m"), minmos=3, minobs=3, timespan=c(1989,2018))
MNMPCAlakes.raw<-fixNamestoLAGOS(MNMPCAlakes.raw,"Minnesota MPCA")
MNMPCAlakes.cln<-cleanAnnualts(MNMPCAlakes.raw)

WIDNRlakes.raw<-makeWIDNRannualts(dir="/Users/jonathanwalter/Box Sync/NSF EAGER Synchrony/Data/LAGOS Extended/WI_DNR",
                                  infovars=c("lake_area_ha", "lake_perim_meters", "nhd_ftype", "nhd_fcode", "hu4_zoneid","hu12_zoneid", 
                                             "state_zoneid", 
                                             "elevation_m"), minmos=3, minobs=3, timespan=c(1989,2018))
WIDNRlakes.raw<-fixNamestoLAGOS(WIDNRlakes.raw, "Wisconsin DNR")
WIDNRlakes.cln<-cleanAnnualts(WIDNRlakes.raw)

#IowaALMlakes<-makeIowaALMannualts(almdat="/Users/jonathanwalter/Box Sync/NSF EAGER Synchrony/Data/Iowa Synchrony Lakes.csv")
